### 2021-02-19 text/draft.tex
reconcilement

For data refinement I use OpenRefine version 3.4.1
(https://openrefine.org/). Data manipulation in OpenRefine is tracked:
the manipulation history of some data can be exported as JSON file and
then reproduced (on the same data on a different machine or on different
data) by loading said file. I will supply a corresponding manipulation
history whenever appropriate. Note that when using OpenRefine with
larger files (and there are many such files in this project), the memory
allocation to OpenRefine must be manually increased (see
https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation
for details).

@@

There were X automatic matches (call them ``suggestions''). The quality
of these suggestions was then evaluated by random sampling. A suggestion
is considered to be correct if and only if the meaning of the keyword
from \texttt{keywords/keywords\_clean\_histogram.json} corresponds to
the meaning of the suggested subject term from Wikidata. Note that for
homonymous or polysemous keywords, it is impossible to confirm a correct
match without further context; those keywords therefore cannot be
reconciled (but see \protect\hyperlink{construction}{section
``Construction''} for a possible solution).

Unfortunately, random sampling showed that the overall quality of the
suggestions was not satisfactory. A two-pronged strategy was adopted to
ameliorate the quality of the reonciliation:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  manual matching of the top results
\item
  automated processes for the rest
\end{enumerate}

!! calculate for each of annif yso mesh how many percent of keywords or
total occurrences are covered by reonciliated data.

>>>>

For data refinement I use OpenRefine version 3.4.1
(https://openrefine.org/). Data manipulation in OpenRefine is tracked:
the manipulation history of some data can be exported as JSON file and
then reproduced (on the same data on a different machine or on different
data) by loading said file. I will supply a corresponding manipulation
history whenever appropriate. Note that when using OpenRefine with
larger files (and there are many such files in this project), the memory
allocation to OpenRefine must be manually increased (see
https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation
for details). ALso note that OpenRefine always counts the number of
records by the first column (a fact which caused me many headaches).

@@

There were X automatic matches (call them ``suggestions''). A suggestion
is correct if and only if the meaning of the keyword from
\texttt{keywords/keywords\_clean\_histogram.json} corresponds to the
meaning of the suggested concept from Wikidata. Note that for homonymous
or polysemous keywords, it is impossible to confirm a correct match
without further context; those keywords therefore cannot be reconciled
(but see \protect\hyperlink{construction}{section ``Construction''} for
a possible solution). Unfortunately, random sampling showed that the
overall quality of the reconciliation was not satisfactory, that is,
there were too many incorrect suggestions. A two-pronged strategy was
adopted to ameliorate the quality of the reconciliation.

First, the suggestions to the top 500 keywords were manually verified.
These keywords account for 14'996 occurrences or 40.638\% of the total
occurrences and are constitute thus an effective lever.

Second, systematic biases were identified and removed. The most
prevalent bias was an due to a suggestion's type. As stated above, the
reconciliation service API was not constrained by type but had access to
the complete Wikidata database. Since Wikidata is an ontology that
encompasses everything, it also features types whose concepts cannot
qualify as subject terms (at least in the present context). The most
prominent example is the type Q13442814 ``scholarly article.'' Wikidata
contains the metadata of many scholarly articles. Now, for some of our
keywords, there is a scholarly article with a title that exactly matches
the keyword; and since there is no restriction concerning the type, the
scholarly article is suggested with high confidence (see
https://github.com/OpenRefine/OpenRefine/wiki/Reconciliation-Service-API
for details). For example, {[}keyword/article{]}. To generalize,
suggestions with types whose concepts are proper names are usually
incorrect. Based on this observation, suggestions with the following
types were rejected: ``scholarly article,'' ``clinical trial,''
``scientific journal,'' ``academic journal,'' ``open access journal,''
``thesis,'' ``doctoral thesis,'' ``natural number.'' Suggestions with
the following types were manually verified (i.e., checked for
correctness): ``human,'' ``album,'' ``film,'' ``musical group,''
``business,'' ``literary work,'' ``television series,''
``organization,'' ``family name,'' ``written work,'' ``video game,''
``single, television series episode,'' ``painting,'' ``commune of
France,'' ``city of the United States,'' ``magazine,'' ``studio album,''
``year,'' ``nonprofit organization,'' ``border town,'' ``international
organization,'' ``political party,'' ``software,'' ``song,''
``website,'' ``article comic strip,'' ``collection,'' ``commune of
Italy,'' ``fictional human,'' ``film,'' ``government agency,''
``village,'' ``academic journal article,'' ``female given name,''
``poem.''

A second bias found was

After completing these steps to ameliorate the quality of the
reconciliation,

then suggests this scholarly article with high confidence.

However, certain types of concepts cannot qualify as subject terms

!! probabilistic random sampling to assess quality, weighted by
occurrences

!! calculate for each of annif yso mesh how many percent of keywords or
total occurrences are covered by reonciliated data: - Qid: 28789
occurrences, 9635 keywords - Yso: 14557 occurrences, 3185 keywords -
Mesh: 19775 occurrences, 4619 keywords
